Supermartingales and Termination Analysis of Statistical PCF

### 2021-04-26 LICS21/confluent.tex
Edit to focus on confluence result for IWC.
Remove everything except the sections on the confluent trace semantics
and syntax. Add a new introduction. Remove references to other parts
of the paper. Generally edit so that it makes sense on its own.

% !TEX root = main.tex
\section{Confluent trace semantics}
\label{sec:confluent}

When proving almost sure termination in this way, it is necessary to consider all of the terms a given term may reduce to, and organise them into a manageable number of cases to assign ranking function values. 
Sometimes however, the reduction that the programmer has in mind may not be strictly the call-by-value order defined so far; 
or the number of cases necessary for defining a sparse ranking function is excessive because the reduction does not proceed in a neat and orderly manner. 
In cases such as these (e.g.~\cref{ex:sum of powers}), it may be much more convenient to instead consider ranking functions with respect to alternative reduction strategies constructed to match the individual term in question, 
thereby reducing the number of cases needed and simplifying the reachable terms.

In order to be able to use alternative reduction strategies to prove properties of the program defined using the usual reduction strategy, we first prove the relation between these different strategies, providing a confluent version of the semantics. It is therefore necessary to have some way of relating the samples that will be used in runs of the program using different reduction strategies. Positions are defined as a way of addressing sample statements within a program independently of the reduction order. Next, a version of the reduction relation is presented that is nondeterministic, so that it allows a choice of what order to perform reductions in. The notion of positions is extended to potential positions, for samples which may appear later in the reduction sequence but not necessarily in the initial term. In order to allow potential positions in different reduction sequences to be considered equivalent, a relation $\sim^*$ is defined, and finally, all of these are used to construct a confluent version of the trace semantics, $\Rightarrow$, that is still nondeterministic in reduction order, but does specify the outcome of random choices. The desired properties of this semantics (confluence and equivalence to the standard semantics) are proved (\Cref{churchRosser,thm:AstEquivalence}), then it is used to give extended versions of ranking functions and the related AST results, \Cref{thm:confluent ranking}.

Non-probabilistic lambda calculi are generally confluent, i.e.~if a term $A$ reduces to both $B_1$ and $B_2$, there is some $C$ 
to which both $B_1$ and $B_2$ reduce,
so the reduction order mostly doesn't matter. 
In the probabilistic case, this may not be true, because $\beta$-reduction can duplicate $\tsample$s, so the outputs of the copies of the sample may be identical or independent, depending on whether the sample is taken before or after $\beta$-reduction. 
There are, however, some restricted variations on the reduction order that do not have this problem.

\medskip
Even with such a restriction, a trace semantics in the style of the one already defined would not be entirely confluent, as, for example, $\red^3(\tsample - \tsample, (1,0,\dots))$ would be either $1$ or $-1$ depending on the order of evaluation of the $\tsample$s, as that determines which sample from the pre-selected sequence is used for each one. To fix this, rather than pre-selecting samples according to the order they'll be drawn in, select them according to the position in the term where they'll be used instead.

@@

Thus we can define a more general reduction relation\footnote{In this section $\to$ always means the relation defined in \cref{def:more general red}.} $\to$.
\begin{definition}
\label{def:more general red}
The binary relation $\to$ is defined by the following rules, each is conditional on a redex occurring at position $\alpha$ in the term $M$:
\begin{align*}
  \text{if } M \mid \alpha = (\lambda x.N) V,\ & M \to M[N[V/x]/\alpha] \\
  \text{if } M \mid \alpha = \underline f (\underline r_1, \dots , \underline r_n),\ & M \to M[\underline{f(r_1,\dots,r_n)}/\alpha] \\
  \text{if } M \mid \alpha = \tY \lambda x. N,\ & M \to M[\lambda z. N[(\tY \lambda x. N)/x] z/\alpha]\\ \text{where }z&\text{ is not free in $N$}\\
  \text{if } M \mid \alpha = \tif{\underline r < 0}{N_1}{N_2},\ & M \to M[N_1/\alpha] \text{ where }r < 0 \\
  \text{if } M \mid \alpha = \tif{\underline r < 0}{N_1}{N_2},\ & M \to M[N_2/\alpha] \text{ where }r \geq 0 \\
  \text{if } M \mid \alpha = \tsample \text{ and $\lambda$ does}&\text{ not occur after $@_2$ or $\tY$ in $\alpha$},\\ & M \to M[\underline r/\alpha] \text{ where } r \in [0,1].
\end{align*}
In each of these cases, $M \mid \alpha$ is the \emph{redex}, and the reduction \emph{takes place at $\alpha$}. Each subterm can be a redex in at most one way, but there can be multiple redexes at different positions.
\end{definition}

@@

%\medskip
Labelling the pre-chosen samples by the positions in the term by using $I^{\{\alpha \mid \, (M \mid \alpha) = \tsample\}}$ as the trace space would not be sufficient to solve the issue of different samples being used in corresponding locations in different reduction sequences because in some cases, a $\tsample$ will be duplicated before being reduced, for example, in $(\lambda x. x \, {\underline 0} \, \mathbin{\underline{+}} \, {x \, \underline 0}) (\lambda y. \tsample)$, both of the $\tsample$ redexes that eventually occur originate at $@_2 ; \lambda$. 
It is therefore necessary to consider possible positions that may occur in other terms reachable from the original term. 
Even this is itself inadequate because some of the positions in different reachable terms need to be considered the same, and the number of reachable terms is in general uncountable, which leads to measure-theoretic issues.

\paragraph*{Reduction relation on skeletons} We are thus led to consider the reduction relation on skeletons (and positions in a skeleton), which can be extended from the definitions on terms in the obvious way, with $\tif{\skeletonPlaceholder < 0}{A}{B}$ reducing nondeterministically to both $A$ and $B$, $\tsample$ reducing to $\skeletonPlaceholder$, and $\skeletonPlaceholder$ considered as (the skeletal equivalent of) a value, so that $(\lambda x.A) \, \skeletonPlaceholder$ reduces to $A[\skeletonPlaceholder / x]$.
For example, we have 
\(
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder}) \, \tsample
\to
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder}) \, \skeletonPlaceholder
\to
\tif{\skeletonPlaceholder < 0}{\skeletonPlaceholder}{\skeletonPlaceholder}
\to
\skeletonPlaceholder
\).

@@

Reduction sequences are used rather than reachable skeletons because if the same skeleton is reached twice, different samples may be needed.

@@

%\medskip
This is still not quite sufficient to attain confluence because sometimes the same samples must be used at corresponding positions in different reduction sequences.

@@

        \item $N \mid \alpha = \tY (\lambda x. A)$, $A$ reduced at $\beta$ is $A'$, $(\gamma_i)_i$ is a list of all the positions where $A' \mid \gamma  = x$, ordered from left to right, $O_1$ is $N$ reduced first at $\alpha;\tY;\lambda;\beta$ then at $\alpha$, and $O_2$ is $N$ reduced first at $\alpha$ then at $\alpha;\lambda;@_1;\gamma_i;Y;\lambda;\beta$ for each $i$ in order where $\gamma_i$ is left of $\beta$ then at $\alpha;\lambda;@_1;\beta$ then at $\alpha;\lambda;@_1;\gamma_i;Y;\lambda;\beta$ for the remaining values of $i$.
    \end{compactenum}
    (in which case $O_1$ and $O_2$ are equal as skeletons, but with different reduction sequences), $O_1'$ and $O_2'$ are the results of applying some reduction sequence to each of $O_1$ and $O_2$ (the same reductions in each case, which is always possible because they're equal skeletons), and $\delta$ is a position in $O_1'$ (or equivalently $O_2'$), then $(O_1',\delta) \sim_c (O_2',\delta)$.
\end{asparaenum}
The $\sim_c$-rules are illustrated in \Cref{sim-diagram}.
\end{definition}

@@

\medskip
Before defining the new version of the reduction relation $\red$, the following lemma is necessary for it to be well-defined.

@@

\begin{definition}[$\Rightarrow$ reduction]
Unlike in the purely call-by-value case, the version of the reduction relation that takes into account samples is still a general relation rather than a function, so it is denoted ``$\Rightarrow$'' instead of ``$\red$'', and it relates $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ to itself.
We write an element of $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ as $(M', s)$ where the \emph{term} $M' \in \Lambda^0$ and $s \in  I^{L_s(M')}$.
\begin{align*}
& (M,s) \Rightarrow (N,s \circ i(M \to N)) \text{ if $M \to N$ at $\alpha$ and either} \\
& \qquad \text{the redex is not $\tsample$, or} \\
& \qquad \text{$M \mid \alpha = \tsample$ and $N = M[\underline{s(\mathit{Sk}(M),\alpha)}/\alpha]$}
\end{align*}
\end{definition}

@@

\paragraph*{Reduction strategy determinising $\Rightarrow$}
The reduction relation $\Rightarrow$ is nondeterministic, so it admits multiple possible reduction strategies. 
%\begin{definition}
A \emph{reduction strategy} starting from a closed term $M$ is a measurable partial function $f$ from $\mathit{Rch}(M)$ to positions, such that for any reachable term $N$ where $f$ is defined, $f(N)$ is a position of a redex in $N$, and if $f(N)$ is not defined, $N$ is a value.
Using a reduction strategy $f$, a subset of $\Rightarrow$ that isn't nondeterministic, $\Rightarrow_f$, can be defined by $(N,s) \Rightarrow_f (N',s')$ just if $(N,s) \Rightarrow (N',s')$ and $N$ reduces to $N'$ with the redex at $f(N)$.
%\end{definition}

@@

With these definitions, it is now possible to relate the confluent trace semantics back to the standard trace semantics.

@@

All of the theorems on the termination of rankable terms therefore extend to other reduction strategies too. The proofs of \Cref{thm:rankable implies termination,thm:partial implies rankable,thm:antitone rankable implies termination,thm:antitone partial implies rankable} are all sufficiently generic with respect to what the reduction relation actually is that they can be directly applied to other reduction strategies. They only require that the number of reductions that can occur without any of them being a $\tY$-reduction is bounded for any starting term (which is true, because \Cref{thm:de groote} applies equally to any reduction strategy). For a reduction strategy $r$ and term $M$, just substitute $N[r(N)]$ being a $\tY$-redex for $N$ being of the form $E[\tY \lambda x. O]$, and $r(N)$ being undefined for $N$ being a value.

The domain of definition of the ranking functions also needs to be changed from the reachable terms $\mathit{Rch}(M)$ to the \emph{reachable terms with respect to the reduction strategy $r$},
\[
\begin{array}{l}
\mathit{Rch}_r(M) := \{N \mid \exists n, (N_i)_{0 \leq i \leq n} : {}\\
\qquad\qquad\qquad N_0 = M, N_n = N, N_i \to N_{i+1} \text{ at } r(N_i)\}.
\end{array}
\]

More explicitly, the modified forms of the theorems are:
\begin{definition}
An \emph{antitone ranking function on $M$ w.r.t.~a reduction strategy $r$} is a measurable function $f:\mathit{Rch}_r(M) \to \mathbb{R}$ such that $f(N) \geq 0$ for all $N$, and there exists an antitone function $\epsilon : \nnReal \to \pReal$ such that
\begin{itemize}
    \item $f(N) \geq \epsilon(f(N)) + f(N')$ if $N[r(N)]$ is a $\tY$-redex and $N$ reduces to $N'$ at $r(N)$
    \item $f(N) \geq \int_I f(N[\underline{x}/r(N)]) \, \Leb(\mathrm{d}x)$ if $N[r(N)] = \tsample$

    \item $f(N) \geq f(N')$ if $r(N)$ is any other redex, where $N \to N'$ at $r(N)$.
\end{itemize}
Any closed term for which an antitone ranking function with respect to a reduction strategy $r$ exists is called \emph{antitone rankable w.r.t.~$r$}. 
\end{definition}

\begin{definition}
An \emph{antitone sparse ranking function on $M$ w.r.t.~a reduction strategy $r$} is a partial function $f : \mathit{Rch}_r(M) \rightharpoonup \mathbb R$ such that there exists some antitone function $\epsilon : \nnReal \to \pReal$ such that
\begin{inparaenum}[(i)]
    \item $f(N) \geq 0$ for all $N \in \dom(f)$,
    \item $M \in \dom(f)$,
    \item for any $N \in \dom(f)$, evaluation of $N$ at the positions specified by $r$ will eventually reach some $O$ such that either $r(O)$ isn't defined or $f(O)$ is, and $f(N) \geq \mathbb E[f(O) + \epsilon(f(O)) \times \text{the number of $\tY$-reduction steps from $N$ to $O$}]$ (where $f(O)$ is taken to be 0 if $O$ is a value outside of $\dom(f)$).
\end{inparaenum}
\iffalse
\begin{itemize}
    \item $f(N) \geq 0$ for all $N$ where $f$ is defined.
    \item $f$ is defined at $M$.
    \item For any $N$ in the domain of definition of $f$, evaluation of $N$ at the positions specified by $r$ will eventually reach some $O$ such that either $r(O)$ isn't defined or $f(O)$ is, and $f(N) \geq \mathbb E[f(O) + \epsilon(f(O)) \times \text{the number of $\tY$-reduction steps from $N$ to $O$}]$ (where $f(O)$ is taken to be 0 if $O$ is a value outside of the domain of $f$).
\end{itemize}
\fi
\end{definition}

If the antitone function is just the constant function $1$ instead, then it is called a \emph{(sparse) ranking function on $M$ w.r.t.~$r$} instead, and the term is called \emph{rankable w.r.t.~$r$}.

\begin{theorem}[Confluent ranking] \label{thm:confluent ranking}
If a closed PPCF term $M$ has a ranking function, sparse ranking function, antitone ranking function or antitone sparse ranking function w.r.t.~a reduction strategy $r$, then $M$ is AST w.r.t.~$r$, and AST.
\end{theorem}

When constructing an (antitone) sparse ranking function with respect to an alternative reduction strategy, it is intuitively simplest to define the sparse ranking function and the reduction strategy together. For each key reachable term where the sparse ranking function is defined, simply decide which sub-term should be reduced next, and how far before the next key reachable term. The only restrictions to bear in mind on evaluating sub-terms are that a $\tsample$ may not be evaluated inside a $\lambda$ that's inside a $\tY$ or on the right of an application, and a $\beta$ redex may not be reduced until its argument is a value. This can be seen as applied to \Cref{ex:sum of powers} at the end of \Cref{apx:confluent}.

\medskip
If, as seems likely, \Cref{thm:antitone rankable implies termination} is complete for terms from which every reachable term is AST, then by the following theorem, \Cref{thm:confluent ranking} can't actually prove the termination of any terms not already provable by \Cref{thm:antitone rankable implies termination} and \Cref{thm:antitone partial implies rankable}.

\begin{restatable}{theorem}{confluentNoStronger}
\label{thm:confluentNoStronger}
For any closed term $M$ and reduction strategy $r$ on $M$, if every term in $\mathit{Rch}_r(M)$ is AST, then every term in $\mathit{Rch}(M)$ is AST.
\end{restatable}

\paragraph*{Custom reduction strategy for AST analysis} However, there are some terms which terminate more simply and directly via some alternative reduction strategy. For these examples, a custom reduction strategy is specified for each term under consideration.

\begin{example}[Sum of powers] \label{ex:sum of powers}
The following term picks a random natural number $n$, then computes $\sum_{k=0}^n k^n$.
\begin{align*}
M  &= (\lambda n. P[n]) \lfloor \tsample^{-1/2} \rfloor \\
P[n]  &= (\lambda p. \Xi[p] \, n) (\lambda x. \Theta[x] \, n)) \\
\Xi[p]  &= \tY \lambda f n. \tif{n = 0}{\underline 0}{p \, n + f (n-1)} \\
\Theta[x]  &= \tY \lambda f n. \tif{n = 0}{\underline 1}{x \times f (n-1)}.
\end{align*}

With the CBV reduction strategy, the recursion in $\Theta[x]$ (that defines the function $k \mapsto k^x$) is executed separately for every term in the sum in $\Xi$. Not only is this much more complicated than simplifying $\Theta[x] \, \underline n$ first, it requires considerably more $\tY$-reduction steps, so many so that $M$ is $\tY$-PAST with respect to a reduction strategy that simplifies $\Theta[x] \, \underline n$ first (call it ``$r$''), but isn't $\tY$-PAST with respect to the standard reduction strategy, so that $M$ is rankable with respect to $r$, but isn't rankable. It is still antitone rankable, but this is much more complicated. Details of the ranking function construction with and without using the confluent semantics are given in \Cref{apx:confluent}.
\end{example}

\begin{remark}
\label{rem:ex corrected using confluent semantics}
Even some of the examples given earlier implicitly use slightly non-standard reduction strategies to make the definitions of their ranking functions slightly simpler. For example, the antitone sparse ranking function in \Cref{ex:unbiased random walk} was defined at $\Theta_2\, \underline n$ (where $\Theta_2 = \tY \, \lambda f n . \, \tif{n = 0}{0}{f \, (n - 1) \oplus_{1/2} f \, (n + 1)}$), but $\Theta_2\, \underline{10}$ actually reduces to $\Theta_2\, (\underline{10} - 1)$ or $\Theta_2\, (\underline{10} + 1)$, and (taking the second of these), it then reduces to $(\lambda z. (\lambda n. \, \tif{n = 0}{0}{\Theta_2 \, (n - 1) \oplus_{1/2} \Theta_2 \, (n + 1)}) z) \, (\underline{10} + 1)$, then $(\lambda z. (\lambda n. \, \tif{n = 0}{0}{\Theta_2 \, (n - 1) \oplus_{1/2} \Theta_2 \, (n + 1)}) z) \, \underline{11}$, bypassing $\Theta_2\, \underline{11}$ entirely. The antitone sparse ranking function in the form given can be justified anyway, by considering it to be using a reduction strategy which is the same as $\cbv$ except that in $\Theta_2\, (\underline f \, \underline n)$, it evaluates the redex at $@_2$ first.

\Cref{ex:higher-order recursion} also makes use of an alternative reduction strategy. In this case, it is to allow the antitone sparse ranking function to skip over all of those terms where the unknown function $F$ is part way through evaluation.
\end{remark}

>>>>

% !TEX root = main.tex
\section{Positions}

@@

Thus we can define a nondeterministic reduction relation $\to$.
\begin{definition}
\label{def:more general red}
The binary relation $\to$ is defined by the following rules, each is conditional on a redex occurring at position $\alpha$ in the term $M$:
\begin{align*}
  \text{if } M \mid \alpha = (\lambda x.N) V,\ & M \to M[N[V/x]/\alpha] \\
  \text{if } M \mid \alpha = \underline f (\underline r_1, \dots , \underline r_n),\ & M \to M[\underline{f(r_1,\dots,r_n)}/\alpha] \\
  \text{if } M \mid \alpha = \tY \lambda x. N,\ & M \to M[\lambda z. N[(\tY \lambda x. N)/x] z/\alpha]\\ \text{where }z&\text{ is not free in $N$}\\
  \text{if } M \mid \alpha = \tif{\underline r < 0}{N_1}{N_2},\ & M \to M[N_1/\alpha] \text{ where }r < 0 \\
  \text{if } M \mid \alpha = \tif{\underline r < 0}{N_1}{N_2},\ & M \to M[N_2/\alpha] \text{ where }r \geq 0 \\
  \text{if } M \mid \alpha = \tsample \text{ and $\lambda$ does}&\text{ not occur after $@_2$ or $\tY$ in $\alpha$},\\ & M \to M[\underline r/\alpha] \text{ where } r \in [0,1].
\end{align*}
In each of these cases, $M \mid \alpha$ is the \emph{redex}, and the reduction \emph{takes place at $\alpha$}. Each subterm can be a redex in at most one way, but there can be multiple redexes at different positions.
\end{definition}

@@

\section{Skeletal Reduction Sequences}
Labelling the pre-chosen samples by the positions in the term by using $I^{\{\alpha \mid \, (M \mid \alpha) = \tsample\}}$ as the trace space would not be sufficient to solve the issue of different samples being used in corresponding locations in different reduction sequences because in some cases, a $\tsample$ will be duplicated before being reduced, for example, in $(\lambda x. x \, {\underline 0} \, \mathbin{\underline{+}} \, {x \, \underline 0}) (\lambda y. \tsample)$, both of the $\tsample$ redexes that eventually occur originate at $@_2 ; \lambda$. 
It is therefore necessary to consider possible positions that may occur in other terms reachable from the original term. 
Even this is itself inadequate because some of the positions in different reachable terms need to be considered the same, and the number of reachable terms is in general uncountable, which leads to measure-theoretic issues.

We are thus led to consider the reduction relation on skeletons. Define a \emph{skeleton} to be a term but, instead of having real constants $\underline r$, it has a placeholder $\skeletonPlaceholder$, so that each term $M$ has a skeleton $\mathit{Sk}(M)$, and each skeleton $S$ can be converted to a term $S[r]$ given a vector $r$ of $n$ real numbers to substitute in, where $n$ is the number of occurrences of $\skeletonPlaceholder$ in $S$. Positions in a skeleton and the reduction relation $\to$ on skeletons can be extended from the definitions on terms in the obvious way, with $\tif{\skeletonPlaceholder < 0}{A}{B}$ reducing nondeterministically to both $A$ and $B$, $\tsample$ reducing to $\skeletonPlaceholder$, and $\skeletonPlaceholder$ considered as (the skeletal equivalent of) a value, so that $(\lambda x.A) \, \skeletonPlaceholder$ reduces to $A[\skeletonPlaceholder / x]$.
For example, we have 
\(
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder}) \, \tsample
\to
(\lambda x . \tif{x < 0}{x}{\skeletonPlaceholder}) \, \skeletonPlaceholder
\to
\tif{\skeletonPlaceholder < 0}{\skeletonPlaceholder}{\skeletonPlaceholder}
\to
\skeletonPlaceholder
\).

@@

Reduction sequences are used rather than reachable skeletons because if the same skeleton is reached twice, different samples may be needed:

@@

\section{Potential Positions}
This is still not quite sufficient to attain confluence because sometimes the same samples must be used at corresponding positions in different reduction sequences.

@@

        \item $N \mid \alpha = \tY (\lambda x. A)$, $A$ reduced at $\beta$ is $A'$, $(\gamma_i)_i$ is a list of all the positions where $A' \mid \gamma  = x$, ordered from left to right, $O_1$ is $N$ reduced first at $\alpha;\tY;\lambda;\beta$ then at $\alpha$, and $O_2$ is $N$ reduced first at $\alpha$ then at $\alpha;\lambda;@_1;\gamma_i;Y;\lambda;\beta$ for each $i$ in order where $\gamma_i$ is left of $\beta$ then at $\alpha;\lambda;@_1;\beta$ then at $\alpha;\lambda;@_1;\gamma_i;Y;\lambda;\beta$ for the remaining values of $i$.
    \end{compactenum}
    (in which case $O_1$ and $O_2$ are equal as skeletons, but with different reduction sequences), $O_1'$ and $O_2'$ are the results of applying some reduction sequence to each of $O_1$ and $O_2$ (the same reductions in each case, which is always possible because they're equal skeletons), and $\delta$ is a position in $O_1'$ (or equivalently $O_2'$), then $(O_1',\delta) \sim_c (O_2',\delta)$.
\end{asparaenum}
\end{definition}

@@

\section{The Confluent Trace Semantics}
Before defining the new version of the reduction relation, the following lemma is necessary for it to be well-defined.

@@

\begin{definition}[$\Rightarrow$ reduction]
This version of the reduction relation now specifies the results of $\tsample$ reductions, but is still nondeterministic with respect to the order of reduction. It relates $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ to itself.
We write an element of $\biguplus_{M \in \Lambda_0} I^{L_s(M)}$ as $(M', s)$ where the term $M' \in \Lambda^0$ and $s \in  I^{L_s(M')}$.
\begin{align*}
& (M,s) \Rightarrow (N,s \circ i(M \to N)) \text{ if $M \to N$ at $\alpha$ and either} \\
& \qquad \text{the redex is not $\tsample$, or} \\
& \qquad \text{$M \mid \alpha = \tsample$ and $N = M[\underline{s(\mathit{Sk}(M),\alpha)}/\alpha]$}
\end{align*}
\end{definition}

@@

In order to show that $\Rightarrow$ behaves as expected, the following lemma is also necessary, in order to show that a sample is never used multiple times in the same reduction sequence:
\begin{lemma} \label{lem:redexDestroyed}
If $M \to N$, with the redex at position $\alpha$, then no position in any term reachable from $N$ is related by $\sim^*$ to $(M,\alpha)$.
\end{lemma}

The reduction relation $\Rightarrow$ is nondeterministic, so it admits multiple possible reduction strategies. 
%\begin{definition}
A \emph{reduction strategy} starting from a closed term $M$ is a measurable partial function $f$ from $\mathit{Rch}(M)$ to positions, such that for any reachable term $N$ where $f$ is defined, $f(N)$ is a position of a redex in $N$, and if $f(N)$ is not defined, $N$ is a value.
Using a reduction strategy $f$, a subset of $\Rightarrow$ that isn't nondeterministic, $\Rightarrow_f$, can be defined by $(N,s) \Rightarrow_f (N',s')$ just if $(N,s) \Rightarrow (N',s')$ and $N$ reduces to $N'$ with the redex at $f(N)$.
%\end{definition}

@@

This reduction strategy allows the confluent trace semantics to be related to the standard version of the trace semantics with a fixed reduction order and linear traces. In \cite{arxiv}, which gives the full definition of the standard trace semantics, this is used to prove the following theorems that allow termination results to be transferred from the confluent trace semantics to the standard trace semantics.

@@
